\textbf{Using the MSU Thesis Class}

### 2022-08-19 docs/msu-thesis.tex
Added usage survey link to docs. Changed docs to use lualatex so that html version can be autogenerated. Added CSS and cfg files for conversion. Moved documentation to docs folder.

% !TEX TS-program = pdflatexmk

@@

\def\msuversion{3.5}
\def\msudate{2022/08/15}
\title{\textbf{Using the MSU Thesis Class}}
\author{\textbf{Alan Munn}\\Department of Linguistics, Languages, and Cultures\\Michigan State University\\\texttt{\href{mailto:amunn@msu.edu}{amunn@msu.edu}}}
\date{Version \msuversion\\\msudate}
\usepackage[T1]{fontenc}
\usepackage[lmargin=.75in,rmargin=.75in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{titling}
\usepackage{array, booktabs, multicol, fancyhdr, xspace,tabularx}
\usepackage{enumitem}
\usepackage{fancyvrb,listings,url}
\usepackage[sf]{titlesec}
\usepackage[section]{placeins}
\usepackage[colorlinks=true]{hyperref}

@@

\IfFileExists{luximono.sty}%
{%
  \usepackage[scaled]{luximono}%
}
{%
  \IfFileExists{beramono.sty}%
  {%
    \usepackage[scaled]{beramono}%
  }{}
}

@@

\newcommand*{\pkg}[1]{\texttt{#1}\xspace}
\newlist{scheme}{itemize}{2}
\setlist[scheme]{label={}}
\setdescription{font={\normalfont}}
\setlength{\droptitle}{-1in}

@@

\abstract{\noindent\begin{quote} This is a class file for MSU theses and dissertations.  It is based on the \pkg{memoir} class, and therefore supports all of the functionality of that class.  It should generate a document which meets all the basic formatting requirements laid out by the \href{https://grad.msu.edu/etd/formatting-guide}{Formatting Guide For Submission of Master’s Theses and Doctoral Dissertations} (March 2020) produced by the Graduate School, including requirements for margins, titlepage, page numbering, section titles, sub- and superscript sizes etc.\end{quote}}

@@

The MSU Thesis Class is based on the \pkg{memoir} document class.  The \pkg{memoir} class is an extensive class that incorporates the functionality of many other packages into it. The class is extensively documented, and the documentation (\pkg{memman.pdf}) should be available with any \TeX\ distribution.  I strongly recommend that you familiarize yourself with \pkg{memoir} as you use the present class.

@@

In addition to any options passed to the \pkg{memoir} class (e.g. |oldfontcommands|; see Section \ref{sec:fonts}, font size), the \pkg{msu-thesis} takes six options: one to specify the type of degree, (see Table~\ref{degrees}), one to enable landscape page numbering, two to enable per-chapter bibliographies and appendices, two to make small adjustments to  Table of Contents.  If no degree option is specified, a Ph.D. dissertation is assumed.

@@

\subsection{The \pkg{[lscape]} option}
The Formatting Guide requires that any lansdscape pages be numbered in landscape mode (i.e. along the long edge of the page) rather than in portrait mode, and rotated in the final PDF. The |[lscape]| option enables this. The option loads both the \pkg{pdflscape} package and \pkg{tikz}, and is implemented as a class option so that you are not required to load those packages if they are not needed.
\subsection{The \pkg{[chapterbib]} option}
This option enables per-chapter bibliographies. Each chapter must have its own bibliography. If you are using \pkg{natbib} for bibliographies, per-chapter bibliographies can be implemented using the \pkg{chapterbib} package. Please read its documentation carefully, as it requires each chapter to be a separate file. A newer alternative which I highly recommend is to use \pkg{biblatex} for your bibliography, which makes per-chapter bibliographies very simple. If you have per-chapter bibliographies you cannot have a global bibliography.
\subsection{The \pkg{[chapterapp]} option}

@@

\section{User commands}
In addition to all the user functionality defined by \pkg{memoir}, the \pkg{msu-thesis} class defines seven new titling commands, and two “appeasement” commands, i.e., commands to tweak the formatting to appease the the thesis office’s inconsistency.
\begin{table}[htpb]
\centering
\begin{tabularx}{.8\textwidth}{>{\ttfamily}lX}
\toprule
{\bs maketitlepage} & Makes the main title page for the thesis.\\
{\bs fieldofstudy\{\}} &  Takes one argument corresponding to your field of study.\\
{\bs dedication\{\}} &  Takes one argument (should be short), your dedication.\\
{\{publicabstract\}} &  Environment for the public abstract (unnumbered, uncounted).\\
{\bs makecopyrightpage} &  Creates the copyright page (use \texttt{\bs makecopyrightpage*} to align the text on the left.)\\
{\bs makededicationpage} &  Creates the dedication page.\\
{\bs makebibliographypage} &  Creates the bibliography cover page.\\
\bottomrule
\end{tabularx}
\caption{Class titling commands}
\end{table}

@@

Most book-length documents are divided into three main kinds of parts: the front matter, the main matter, and the back matter.  In \pkg{memoir} (and therefore \pkg{msu-thesis}) these sections are preceded by the commands |\frontmatter|, |\mainmatter|, and |\backmatter|.  These commands tell the class when to change the page numbering, for example, front matter pages  use lower case roman numerals, but main matter pages use arabic numerals.

The basic outline of a dissertation or thesis document is shown in Table~\ref{structure}: (your thesis may not have all of these parts, but if it does, they should be in this order.) You can find an annotated copy of the basic structure in the \href{https://ctan.org/tex-archive/macros/latex/contrib/msu-thesis/samples}{\pkg{samples}} folder of the \pkg{msu-thesis} documentation.

\begin{table}[ht]
\begin{scheme}
\item |\begin{document}|
\item |\frontmatter|
\begin{scheme}
	\item Titlepage
	\item Public Abstract
	\item Abstract
	\item Copyright Page
	\item Dedication
	\item Acknowledgements
	\item Preface
	\item Table of Contents
	\item List of Tables
	\item List of Figures
	\item List of Abbreviations/Symbols
\end{scheme}
\item |\mainmatter|
\begin{scheme}
	\item Chapter 1
	\item Chapter 2
	\item \ldots
	\item Chapter $n$
	\item{Appendices}
\end{scheme}
\item |\backmatter|
\begin{scheme}
	\item{Bibliography}
\end{scheme}
\item |\end{document}|
\end{scheme}
\caption{Basic structure of a thesis or dissertation}\label{structure}
\end{table}
\subsection{What does the class file do?}
The class itself sets up the margins, page numbering, and formatting of all of the required pieces.  It doesn't put everything in the right order, so you are responsible for that.  The best way to do this is to use the included sample \LaTeX\ file as a basic template for your thesis.
\subsection{Line Spacing}
The \pkg{memoir} class provides commands for single and double spacing (|\SingleSpacing| and |\DoubleSpacing|) most of your thesis will be double spaced, but certain parts (such as the table of contents, lists of figures/tales, bibliography) may be single spaced. This duplicates the functionality of the \pkg{setspace} package, which therefore should not be used.  For bibliography spacing issues, see also Section \ref{bibspacing}.

\subsection{Font changing commands}\label{sec:fonts}
\sloppy The \pkg{memoir} class enforces the so-called ``new'' font changing commands (although they have now been in use for many years.) This means that it does not allow you to use the commands |\bf|, |\it|, |\rm|, etc.  Unfortunately there are some old packages that still use these commands, and there is also some outdated information on the web that might have examples using these commands. You should generally not use these commands (see \href{http://www.tex.ac.uk/cgi-bin/texfaq2html?label=2letterfontcmd}{Two Letter Font Commands}); if you do encounter problems, you can pass the option |[oldfontcommands]| to the class.
\subsection{Extra Packages}
The \pkg{memoir} class replicates the functionality of many independent packages, so if you are transferring \LaTeX\ source from other documents using the \pkg{article} class, you may encounter some warnings or errors.  The \pkg{memoir} documentation has substantial information about the packages it emulates, and the packages for which it provides equivalent functionality.  See Chap. 18 Sec. 24 of the \pkg{memoir} manual for details.  Quoting from the manual:
\begin{quote} At the time of writing the emulated packages are: {\sffamily abstract, appendix, booktabs, ccaption, chngcntr, crop, enumerate, epigraph, ifmtarg, ifpdf, index, makeidx, moreverb, needspace, newfile, nextpage, pagenote, patchcmd, parskip, setspace, shortvrb, showidx, titleref, tocbibind, tocloft, verbatim, and verse.} As well as the emulated packages memoir provides functions equivalent to those in the following packages, although the class does not prevent you from using them: {\sffamily fancyhdr, framed, geometry, sidecap, subfigure, and titlesec.}\\\hfill{\normalfont\emph{The Memoir Class} 10th impression, 2015, p.345}
\end{quote}
\subsection{Captions}
\subsubsection{Caption delimiter}\label{captiondelim}
Because the thesis office doesn't understand anything about punctuation, it requires the colon that would appear in a caption to also appear in the contents lists. Since they don't care about the colon itself, but just consistency I've removed colons from the captions entirely, since leaving it in makes the contents lists look really stupid. But in case you or your advisor want the colon in captions, you can add the colon back in by adding   |\renewcommand{\msucaptiondelim}{:\space}| to your preamble, and the class will make sure the Lists of Tables and Figures match in formatting. If you have other floating objects with captions (e.g., algorithms, listings) you may need to change their delimiter too (see below). How you do this may depend on the particular packages you are using.
\subsubsection{New lists}
If you are using other kinds of floats such as listings, you should use the Memoir |\newlistof| command to make the relevant list. The class provides a macro |\msucaptiondelim| which can be used to make sure the formatting of the captions in the list-of-X match the setting you have for Figures and Tables.  Here is an example of how you would add a listings list. For lists of algorithms, see section \ref{algorithms} below. The following code goes in your preamble:
\begin{quote}
\begin{lstlisting}
\newlistof{listoflistings}{lol}{List of Listings}
\newfloat[chapter]{listing}{lol}{Listing}
\newlistentry{listing}{lol}{0}
\renewcommand*{\cftlistingname}{Listing\space}
\renewcommand*{\cftlistingaftersnum}{\msucaptiondelim}
\end{lstlisting}
\end{quote}

@@

\subsection{Algorithm packages\label{algorithms}}
Because the various algorithm formatting packages such as \pkg{algorithm2e} and \pkg{algorithmicx} (or its related packages) sometimes create incompatibilities with the \pkg{memoir} float methods, the class provides you with floating |{algorithm}| environment and the corresponding list of algorithms, so you do not have to create your own list of algorithms. Here are some specific important points about using \pkg{algorithm2e} and \pkg{algorithmicx}.
\begin{itemize}[label=\textbullet]
\item Do \emph{not} load the \pkg{algorithm} package.
\item If you use the \pkg{algorithm2e} package you \emph{must} load it with the |[algo2e]| option (along with any other options you need).
\end{itemize}
\subsubsection{\pkg{algorithmicx} package} If you are using the \pkg{algorithmicx} package, any of its related packages (including its older predecessor \pkg{algorithmic}) which suggest that you use the \pkg{algorithm} package to create floating algorithms, you can load the \pkg{float} package to use the its float styling capabilities instead of loading the \pkg{algorithm} package. You should \emph{not} however use the \pkg{float} package to create new floating environments. You must use the \pkg{memoir} methods to create new floating environments and only then load the \pkg{float} package.

If you would normally load the \pkg{algorithm} package to create floating algorithms do the following instead. If you are creating any other new floating environments (like listings as described above) you \emph{must} create these environments first \emph{before} loading the \pkg{float} package otherwise your document will not compile.
\begin{quote}
\begin{lstlisting}
% This code should only appear AFTER you have created any other new lists
\usepackage{float}
\floatstyle{ruled}% or any of the predefined float styles
\restylefloat{algorithm}
\end{lstlisting}
\end{quote}
\subsubsection{Algorithm caption delimiter}
If you have changed the |\msucaptiondelimeter| to use a colon (see \ref{captiondelim}) then depending on the particular algorithm package you are using, there may be a mismatch between the format of the caption in the algorithm itself and in the List of Algorithms. The Graduate School thinks the presence or absence of a colon in captions and lists is of vital importance, and they may complain about this. Here are two options to use if they complain:

@@

\subsection{The \pkg{microtype} package and contents lists}
If you use the \pkg{microtype} package with |protrusion| set, it’s best to turn it off before your contents lists.  You can do this by using |\microtypesetup{protrusion=false}|
before the contents lists and |\microtypesetup{protrusion=true}| afterwards.  Schematically:

@@

\subsubsection{Public abstract}
If your thesis requires a public abstract, it should go immediately before the regular abstract. Use the |{publicabstract}| environment for this. Please note that you \emph{must} have an abstract even if you also have a public abstract.
\subsection{Landscape figures and tables}
If you have large figures and tables that must be rotated, you should use the |[lscape]| class option.  This enables a pagestyle |lscape| and places the page numbers correctly on the long edge of the page. The option loads the \pkg{pdflscape} package which provides a |landscape| environment to place the landscape figure in. Before the landscape environment, you need to issue a |\clearpage| command and change the pagestyle to |lscape|, and afterwards, change it back to |plain|. Schematically, then, any landscape pages should be created in the following way.

@@

\subsubsection{Appendices per chapter}
If you need to add an appendix or appendices to a chapter, you must use the \pkg{[chapterapp]} class option. The code for the appendices themselves is  exactly as described above. This means that the appendix itself is introduced by the |\chapter| command, \emph{not} a |\section| command (because the this seems to be how the thesis office wants things to look).  Furthermore,  per-chapter appendices should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

\subsection{Bibliographies}
The class places no constraints on how you do your bibliography. It doesn't care whether you are using the \pkg{chapterbib} package with \pkg{natbib} or the per-chapter bibliography functionality provided by \pkg{biblatex}. If you have per-chapter bibliographies, however, you \emph{must} use the |[chapterbib]| class option.  This option lets the class know that you are using per-chapter bibliographies.   Before any bibliography (whether global or per-chapter) you should use the |\makebibliographypage| command. Note that per-chapter bibliographies should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

\subsubsection{Bibliography spacing}\label{bibspacing}
Please note that although bibliography entries must be single spaced, there must be a space between each entry. This parameter is set by your bibliography package and not by the class itself. If you are using \pkg{natbib} or \pkg{biblatex} these spacing parameters are set for you by the class. If you are using another package for your bibliography you are responsible for adjusting the relevant spacing parameters.

@@

\section{Bugs and redistribution}
\subsection{Current version}
The latest copy of \pkg{msu-thesis} will always be available on  \href{https://ctan.org/pkg/msu-thesis?lang=en}{CTAN} and on \href{https://github.com/amunn/msu-thesis}{GitHub}. If you use MiKTeX, or a current year TeX Live then their respective package managers can update or install the latest version. If you are using Overleaf, you may not have the most recent version, since Overleaf is usually about one year behind in its TeX distribution.  If you are using a Linux-based distribution, or your TeXLive distribution is not the most recent year, you may need to install the latest copy in your local |texmf| directory or simply in the same directory as your main source file. If you want others to use the class, \textsc{please} point them to CTAN, or tell them to use an up-to-date TeX distribution so that they can use an up-to-date version with all accompanying documentation and examples rather than passing along your copy. Please report \textsc{any} problems you have with the class to me \href{mailto:amunn@msu.edu}{amunn@msu.edu}, as this will aid in making things easier for those who follow you. You may also report actual bugs or class-caused formatting issues using GitHub: \href{https://github.com/amunn/msu-thesis}{github.com/amunn/msu-thesis}, but in my experience, most issues are more easily dealt with using email.  The Graduate School rarely announces changes it makes to the formatting requirements, and it is often very inconsistent in what kinds of things it chooses to get annoyed about, so your problem reports are extremely helpful for keeping the class in compliance with the latest standards.

@@

Since the current version of the \pkg{msu-thesis} class satisfies the Graduate School requirements at any one time, as these requirements change, you may find that if you need to recompile your thesis after you have graduated that the formatting changes. To avoid this, I would recommend saving an archived version of the \pkg{msu-thesis.cls} file in the same folder as your thesis file.  You only need to do this once you have completely finished your thesis however; there's no need to do it during the writing process. 

@@

Previous versions of \pkg{msu-thesis} (versions 1.0--1.4b) were created for non-electronic submission and should not be used. Versions 1.5--1.6 (the first electronic submission versions) contained a |\makeabstract| command and a separate |msuabstract| environment. These have been removed as of version 1.7.  Users who are upgrading from earlier versions to version 1.7 do not need to change these commands however but their use will trigger a warning.) Version 2.0 implemented landscape pages; version 2.1 added support for \pkg{hyperref}. Versions 2.2--2.5 mainly accommodated to the ever changing whims of the MSU Graduate School. Version 2.6 fixed some appendix bugs introduced by the previous whim accommodation. Version 2.7 fixed various small spacing issues, fixed a bug which prevented a new chapter page to be a landscape page, added support for the \pkg{apacite} package, and updated degree options. Version 2.8 fixed some spacing and capitalization issues and added support for the public abstract. Version 2.9 removed subscript/superscript size restrictions, added font size options and made small spacing changes to appease the thesis office. Version 3.0 added full support for per-chapter bibliographies and appendices, and automated bibliography spacing for |natbib| and |biblatex| produced bibliographies. Version 3.1 added support for an |{algorithm}| environment and a list of algorithms to combat incompatibilities with the \pkg{algorithm} and \pkg{algorithm2e} packages.  Version 3.2 fixed some lingering problems with per-chapter appendices. Version 3.3 added code to accommodate the insane inconsistencies in formatting requirements that the thesis office requires, including a new contents option which hopefully should never be used, and fixed some numbering problems with per-chapter appendices. Version 3.4 added a starred option to the |\makecopyrightpage| command after complaints about placement. The documentation has also been updated to include more information about caption formatting and removed colons from the default caption format. Version 3.5 added user level way to change the spacing after the author’s name in the abstract, to deal with thesis office inconsistency.

>>>>

% !TEX TS-program = lualatexmk

@@

\def\msuversion{3.5}
\def\msudate{2022/08/15}
\title{\textbf{Using the MSU Thesis Class}}
\author{\textbf{Alan Munn}\\Department of Linguistics, Languages, and Cultures\\Michigan State University\\\texttt{\href{mailto:amunn@msu.edu}{amunn@msu.edu}}}
\date{Version \msuversion\\\msudate}
\usepackage{fontspec}
\usepackage[lmargin=.75in,rmargin=.75in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{titling}
\usepackage{array, booktabs, multicol, fancyhdr, xspace,tabularx}
\usepackage{enumitem}
\usepackage{fancyvrb,listings,url}
\usepackage[sf,small]{titlesec}
\usepackage[section]{placeins}
\usepackage[colorlinks=true]{hyperref}
\setmonofont{Inconsolatazi4}

@@

@@

\newcommand*{\pkg}[1]{\texttt{#1}}
\newlist{scheme}{itemize}{2}
\setlist[scheme]{label={}}
\setdescription{font={\normalfont}}
\setlength{\droptitle}{-1in}

@@

\abstract{\noindent\begin{quote} This is a class file for MSU theses and dissertations.  It is based on the \pkg{memoir}\ class, and therefore supports all of the functionality of that class.  It should generate a document which meets all the basic formatting requirements laid out by the \href{https://grad.msu.edu/etd/formatting-guide}{Formatting Guide For Submission of Master’s Theses and Doctoral Dissertations}\ (March 2020) produced by the Graduate School, including requirements for margins, titlepage, page numbering, section titles, sub- and superscript sizes etc.\end{quote}}

@@

The MSU Thesis Class is based on the \pkg{memoir}\ document class.  The \pkg{memoir}\ class is an extensive class that incorporates the functionality of many other packages into it. The class is extensively documented, and the documentation (\pkg{memman.pdf}) should be available with any \TeX\ distribution.  I strongly recommend that you familiarize yourself with \pkg{memoir}\ as you use the present class.

@@

In addition to any options passed to the \pkg{memoir}\ class (e.g. |oldfontcommands|; see Section \ref{sec:fonts}, font size), the \pkg{msu-thesis}\ takes six options: one to specify the type of degree, (see Table~\ref{degrees}), one to enable landscape page numbering, two to enable per-chapter bibliographies and appendices, two to make small adjustments to  Table of Contents.  If no degree option is specified, a Ph.D. dissertation is assumed.

@@

\subsection{The \pkg{[lscape]}~option}
The Formatting Guide requires that any lansdscape pages be numbered in landscape mode (i.e. along the long edge of the page) rather than in portrait mode, and rotated in the final PDF. The |[lscape]| option enables this. The option loads both the \pkg{pdflscape}\ package and \pkg{tikz}, and is implemented as a class option so that you are not required to load those packages if they are not needed.
\subsection{The \pkg{[chapterbib]}~option}
This option enables per-chapter bibliographies. Each chapter must have its own bibliography. If you are using \pkg{natbib}\ for bibliographies, per-chapter bibliographies can be implemented using the \pkg{chapterbib}\ package. Please read its documentation carefully, as it requires each chapter to be a separate file. A newer alternative which I highly recommend is to use \pkg{biblatex}\ for your bibliography, which makes per-chapter bibliographies very simple. If you have per-chapter bibliographies you cannot have a global bibliography.
\subsection{The \pkg{[chapterapp]}~option}

@@

\section{User commands}
In addition to all the user functionality defined by \pkg{memoir}, the \pkg{msu-thesis}\ class defines seven new titling commands, and two “appeasement” commands, i.e., commands to tweak the formatting to appease the the thesis office’s inconsistency.
\begin{table}[htpb]
\centering
\begin{tabularx}{.8\textwidth}{>{\ttfamily}lX}
\toprule
{\bs maketitlepage} & Makes the main title page for the thesis.\\
{\bs fieldofstudy\{\}} &  Takes one argument corresponding to your field of study.\\
{\bs dedication\{\}} &  Takes one argument (should be short), your dedication.\\
{\{publicabstract\}} &  Environment for the public abstract (unnumbered, uncounted).\\
{\bs makecopyrightpage} &  Creates the copyright page (use \texttt{\bs makecopyrightpage*} to align the text on the left.)\\
{\bs makededicationpage} &  Creates the dedication page.\\
{\bs makebibliographypage} &  Creates the bibliography cover page.\\
\bottomrule
\end{tabularx}
\caption{Class titling commands}
\end{table}

@@

Most book-length documents are divided into three main kinds of parts: the front matter, the main matter, and the back matter.  In \pkg{memoir}\ (and therefore \pkg{msu-thesis}) these sections are preceded by the commands |\frontmatter|, |\mainmatter|, and |\backmatter|.  These commands tell the class when to change the page numbering, for example, front matter pages  use lower case roman numerals, but main matter pages use arabic numerals.

The basic outline of a dissertation or thesis document is shown in Table~\ref{structure}: (your thesis may not have all of these parts, but if it does, they should be in this order.) You can find an annotated copy of the basic structure in the \href{https://ctan.org/tex-archive/macros/latex/contrib/msu-thesis/samples}{\pkg{samples}}\ folder of the \pkg{msu-thesis}\ documentation.

\begin{table}[ht]
\begin{scheme}
\item |\begin{document}|
\item |\frontmatter|
\begin{scheme}
	\item Titlepage
	\item Public Abstract
	\item Abstract
	\item Copyright Page
	\item Dedication
	\item Acknowledgements
	\item Preface
	\item Table of Contents
	\item List of Tables
	\item List of Figures
	\item List of Abbreviations/Symbols
\end{scheme}
\item |\mainmatter|
\begin{scheme}
	\item Chapter 1
	\item Chapter 2
	\item \ldots
	\item Chapter $n$
	\item{Appendices}
\end{scheme}
\item |\backmatter|
\begin{scheme}
	\item{Bibliography}
\end{scheme}
\item |\end{document}|
\end{scheme}
\caption{Basic structure of a thesis or dissertation}\label{structure}
\end{table}
\subsection{What does the class file do?}
The class itself sets up the margins, page numbering, and formatting of all of the required pieces.  It doesn't put everything in the right order, so you are responsible for that.  The best way to do this is to use the included sample \LaTeX\ file as a basic template for your thesis.
\subsection{Line Spacing}
The \pkg{memoir}\ class provides commands for single and double spacing (|\SingleSpacing| and |\DoubleSpacing|) most of your thesis will be double spaced, but certain parts (such as the table of contents, lists of figures/tales, bibliography) may be single spaced. This duplicates the functionality of the \pkg{setspace}\ package, which therefore should not be used.  For bibliography spacing issues, see also Section \ref{bibspacing}.

\subsection{Font changing commands}\label{sec:fonts}
\sloppy The \pkg{memoir}\ class enforces the so-called ``new'' font changing commands (although they have now been in use for many years.) This means that it does not allow you to use the commands |\bf|, |\it|, |\rm|, etc.  Unfortunately there are some old packages that still use these commands, and there is also some outdated information on the web that might have examples using these commands. You should generally not use these commands (see \href{http://www.tex.ac.uk/cgi-bin/texfaq2html?label=2letterfontcmd}{Two Letter Font Commands}); if you do encounter problems, you can pass the option |[oldfontcommands]| to the class.
\subsection{Extra Packages}
The \pkg{memoir}\ class replicates the functionality of many independent packages, so if you are transferring \LaTeX\ source from other documents using the \pkg{article}\ class, you may encounter some warnings or errors.  The \pkg{memoir}\ documentation has substantial information about the packages it emulates, and the packages for which it provides equivalent functionality.  See Chap. 18 Sec. 24 of the \pkg{memoir}\ manual for details.  Quoting from the manual:
\begin{quote} At the time of writing the emulated packages are: {\sffamily abstract, appendix, booktabs, ccaption, chngcntr, crop, enumerate, epigraph, ifmtarg, ifpdf, index, makeidx, moreverb, needspace, newfile, nextpage, pagenote, patchcmd, parskip, setspace, shortvrb, showidx, titleref, tocbibind, tocloft, verbatim, and verse.} As well as the emulated packages memoir provides functions equivalent to those in the following packages, although the class does not prevent you from using them: {\sffamily fancyhdr, framed, geometry, sidecap, subfigure, and titlesec.}\\\hfill{\normalfont\emph{The Memoir Class} 10th impression, 2015, p.345}
\end{quote}
\subsection{Captions}
\subsubsection{Caption delimiter}\label{captiondelim}
Because the thesis office doesn't understand anything about punctuation, it requires the colon that would appear in a caption to also appear in the contents lists. Since they don't care about the colon itself, but just consistency I've removed colons from the captions entirely, since leaving it in makes the contents lists look really stupid. But in case you or your advisor want the colon in captions, you can add the colon back in by adding   |\renewcommand{\msucaptiondelim}{:\space}| to your preamble, and the class will make sure the Lists of Tables and Figures match in formatting. If you have other floating objects with captions (e.g., algorithms, listings) you may need to change their delimiter too (see below). How you do this may depend on the particular packages you are using.
\subsubsection{New lists}
If you are using other kinds of floats such as listings, you should use the Memoir |\newlistof| command to make the relevant list. The class provides a macro |\msucaptiondelim| which can be used to make sure the formatting of the captions in the list-of-X match the setting you have for Figures and Tables.  Here is an example of how you would add a listings list. For lists of algorithms, see section \ref{algorithms} below. The following code goes in your preamble:
\begin{quote}
\begin{lstlisting}
\newlistof{listoflistings}{lol}{List of Listings}
\newfloat[chapter]{listing}{lol}{Listing}
\newlistentry{listing}{lol}{0}
\renewcommand*{\cftlistingname}{Listing\space}
\renewcommand*{\cftlistingaftersnum}{\msucaptiondelim}
\end{lstlisting}
\end{quote}

@@

\subsection{Algorithm packages\label{algorithms}}
Because the various algorithm formatting packages such as \pkg{algorithm2e}\ and \pkg{algorithmicx}\ (or its related packages) sometimes create incompatibilities with the \pkg{memoir}\ float methods, the class provides you with floating |{algorithm}| environment and the corresponding list of algorithms, so you do not have to create your own list of algorithms. Here are some specific important points about using \pkg{algorithm2e}\ and \pkg{algorithmicx}.
\begin{itemize}[label=\textbullet]
\item Do \emph{not} load the \pkg{algorithm}\ package.
\item If you use the \pkg{algorithm2e}\ package you \emph{must} load it with the |[algo2e]| option (along with any other options you need).
\end{itemize}
\subsubsection{\pkg{algorithmicx}~package} If you are using the \pkg{algorithmicx}\ package, any of its related packages (including its older predecessor \pkg{algorithmic}) which suggest that you use the \pkg{algorithm}\ package to create floating algorithms, you can load the \pkg{float}\ package to use the its float styling capabilities instead of loading the \pkg{algorithm}\ package. You should \emph{not} however use the \pkg{float}\ package to create new floating environments. You must use the \pkg{memoir}\ methods to create new floating environments and only then load the \pkg{float}\ package.

If you would normally load the \pkg{algorithm}\ package to create floating algorithms do the following instead. If you are creating any other new floating environments (like listings as described above) you \emph{must} create these environments first \emph{before} loading the \pkg{float}\ package otherwise your document will not compile.
\begin{quote}
\begin{lstlisting}
% This code should only appear AFTER you have created any other new lists
\usepackage{float}
\floatstyle{ruled}% or any of the predefined float styles
\restylefloat{algorithm}
\end{lstlisting}
\end{quote}
\subsubsection{Algorithm caption delimiter}
If you have changed the |\msucaptiondelimeter| to use a colon (see \ref{captiondelim}) then depending on the particular algorithm package you are using, there may be a mismatch between the format of the caption in the algorithm itself and in the List of Algorithms. The Graduate School thinks the presence or absence of a colon in captions and lists is of vital importance, and they may complain about this. Here are two options to use if they complain:

@@

\subsection{The \pkg{microtype}~package and contents lists}
If you use the \pkg{microtype}\ package with |protrusion| set, it’s best to turn it off before your contents lists.  You can do this by using |\microtypesetup{protrusion=false}|
before the contents lists and |\microtypesetup{protrusion=true}| afterwards.  Schematically:

@@

\subsubsection{Public abstract}
If your thesis requires a public abstract, it should go immediately before the regular abstract. Use the |{publicabstract}| environment for this. Please note that you \emph{must} have an abstract even if you also have a public abstract.
\subsection{Landscape figures and tables}
If you have large figures and tables that must be rotated, you should use the |[lscape]| class option.  This enables a pagestyle |lscape| and places the page numbers correctly on the long edge of the page. The option loads the \pkg{pdflscape}\ package which provides a |landscape| environment to place the landscape figure in. Before the landscape environment, you need to issue a |\clearpage| command and change the pagestyle to |lscape|, and afterwards, change it back to |plain|. Schematically, then, any landscape pages should be created in the following way.

@@

\subsubsection{Appendices per chapter}
If you need to add an appendix or appendices to a chapter, you must use the \pkg{[chapterapp]}\ class option. The code for the appendices themselves is  exactly as described above. This means that the appendix itself is introduced by the |\chapter| command, \emph{not} a |\section| command (because the this seems to be how the thesis office wants things to look).  Furthermore,  per-chapter appendices should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

\subsection{Bibliographies}
The class places no constraints on how you do your bibliography. It doesn't care whether you are using the \pkg{chapterbib}\ package with \pkg{natbib}\ or the per-chapter bibliography functionality provided by \pkg{biblatex}. If you have per-chapter bibliographies, however, you \emph{must}\ use the |[chapterbib]| class option.  This option lets the class know that you are using per-chapter bibliographies.   Before any bibliography (whether global or per-chapter) you should use the |\makebibliographypage| command. Note that per-chapter bibliographies should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

\subsubsection{Bibliography spacing}\label{bibspacing}
Please note that although bibliography entries must be single spaced, there must be a space between each entry. This parameter is set by your bibliography package and not by the class itself. If you are using \pkg{natbib}\ or \pkg{biblatex}\ these spacing parameters are set for you by the class. If you are using another package for your bibliography you are responsible for adjusting the relevant spacing parameters.

@@

\section{Bugs and redistribution}
\subsection{Current version}
The latest copy of \pkg{msu-thesis}\ will always be available on  \href{https://ctan.org/pkg/msu-thesis?lang=en}{CTAN} and on \href{https://github.com/amunn/msu-thesis}{GitHub}. If you use MiKTeX, or a current year TeX Live then their respective package managers can update or install the latest version. If you are using Overleaf, you may not have the most recent version, since Overleaf is usually about one year behind in its TeX distribution.  If you are using a Linux-based distribution, or your TeXLive distribution is not the most recent year, you may need to install the latest copy in your local |texmf| directory or simply in the same directory as your main source file. If you want others to use the class, \textsc{please} point them to CTAN, or tell them to use an up-to-date TeX distribution so that they can use an up-to-date version with all accompanying documentation and examples rather than passing along your copy. Please report \textsc{any} problems you have with the class to me \href{mailto:amunn@msu.edu}{amunn@msu.edu}, as this will aid in making things easier for those who follow you. You may also report actual bugs or class-caused formatting issues using GitHub: \href{https://github.com/amunn/msu-thesis}{github.com/amunn/msu-thesis}, but in my experience, most issues are more easily dealt with using email.  The Graduate School rarely announces changes it makes to the formatting requirements, and it is often very inconsistent in what kinds of things it chooses to get annoyed about, so your problem reports are extremely helpful for keeping the class in compliance with the latest standards.
\subsection{Usage survey}
If you are using the class, please take a moment to fill out my \href{https://forms.gle/bm2AUzthFxW2naK36}{anonymous usage survey}. Thanks! You should be signed into your MSU Google account to access the form, but the form itself is anonymous.

@@

Since the current version of the \pkg{msu-thesis}\ class satisfies the Graduate School requirements at any one time, as these requirements change, you may find that if you need to recompile your thesis after you have graduated that the formatting changes. To avoid this, I would recommend saving an archived version of the \pkg{msu-thesis.cls}\ file in the same folder as your thesis file.  You only need to do this once you have completely finished your thesis however; there's no need to do it during the writing process. 

@@

Previous versions of \pkg{msu-thesis}\ (versions 1.0--1.4b) were created for non-electronic submission and should not be used. Versions 1.5--1.6 (the first electronic submission versions) contained a |\makeabstract| command and a separate |msuabstract| environment. These have been removed as of version 1.7.  Users who are upgrading from earlier versions to version 1.7 do not need to change these commands however but their use will trigger a warning.) Version 2.0 implemented landscape pages; version 2.1 added support for \pkg{hyperref}. Versions 2.2--2.5 mainly accommodated to the ever changing whims of the MSU Graduate School. Version 2.6 fixed some appendix bugs introduced by the previous whim accommodation. Version 2.7 fixed various small spacing issues, fixed a bug which prevented a new chapter page to be a landscape page, added support for the \pkg{apacite}\ package, and updated degree options. Version 2.8 fixed some spacing and capitalization issues and added support for the public abstract. Version 2.9 removed subscript/superscript size restrictions, added font size options and made small spacing changes to appease the thesis office. Version 3.0 added full support for per-chapter bibliographies and appendices, and automated bibliography spacing for |natbib| and |biblatex| produced bibliographies. Version 3.1 added support for an |{algorithm}| environment and a list of algorithms to combat incompatibilities with the \pkg{algorithm}\ and \pkg{algorithm2e}\ packages.  Version 3.2 fixed some lingering problems with per-chapter appendices. Version 3.3 added code to accommodate the insane inconsistencies in formatting requirements that the thesis office requires, including a new contents option which hopefully should never be used, and fixed some numbering problems with per-chapter appendices. Version 3.4 added a starred option to the |\makecopyrightpage| command after complaints about placement. The documentation has also been updated to include more information about caption formatting and removed colons from the default caption format. Version 3.5 added user level way to change the spacing after the author’s name in the abstract, to deal with thesis office inconsistency.
